---
title: "college_majors_second_attempt"
author: "ASG"
date: "7/3/2021"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(scales)
library(skimr)
library(moderndive)
```
reading in the data
```{r}
recentGrads <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-10-16/recent-grads.csv")
```

viewing raw data
```{r}
glimpse(recentGrads)
slice_sample(recentGrads,n = 5)
```

selecting variables and running summary statistics
```{r}
grads <- recentGrads |> 
  select(Major, Major_category, total_with_major = Total, sample_size_FT_year_round = Sample_size, Men, Women, Employed, Unemployed, median_earnings_FT = Median)
grads |> skim()
```

From the raw data we see that some data is missing / incomplete and some samples for FT, year round earnings are very small. We try to find what is missing or has small sample sizes

```{r}
small_n <- grads |> filter(between(sample_size_FT_year_round,0,40)) |> arrange(sample_size_FT_year_round)
```

The 'food science' major has NA values for 'total_with_major', Men and Women variables.
About 25% or 44/173 majors have a small sample size (< 40) for median earnings of full time, year round graduates.

As exploratory analysis, we could just group the majors into categories and examine the median of median FT earnings within each major category
```{r}
major_categories <- grads |> 
  group_by(Major_category) |> 
  summarize(majors_in_category = n(),
            median_of_median_FT_earnings = median(median_earnings_FT),
            total_grads_category = sum(total_with_major, na.rm = TRUE),
            sample_size_FT = sum(sample_size_FT_year_round, na.rm = TRUE),
            men_in_category = sum(Men, na.rm = TRUE),
            women_in_category = sum(Women, na.rm = TRUE),
            fracn_unemployed_in_category =
            sum(Unemployed, na.rm=TRUE) / sum(Employed, na.rm = TRUE)
            )
```

Visualizing the major_categories data
```{r}
major_categories |> 
  mutate(Major_category = fct_reorder(Major_category, median_of_median_FT_earnings)) |> 
  ggplot(aes(x = Major_category, y = median_of_median_FT_earnings)) + 
  geom_col(aes(fill = fracn_unemployed_in_category)) + 
  coord_flip() +
  labs(x = "",
       y = "median of median salaries in category") +
  scale_y_continuous(labels = dollar_format()) + 
  theme(legend.position = "none")
```
