---
title: "collegemajors_salaries"
author: "ASG"
date: "5/26/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(scales)
```

```{r}
recentGrads <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-10-16/recent-grads.csv")
```

### Some common data processing for next few queries

```{r}
majrsProcessed <- recentGrads  %>% 
  mutate(Major = str_to_title(Major))
```

### What are the major fields / categories by popularity?

```{r}
majrsProcessed %>% 
  count(Major_category, wt = Total, sort = TRUE) %>% 
  mutate(Major_category = fct_reorder(Major_category, n)) %>% 
  ggplot(aes(Major_category,n, fill = Major_category)) +
  geom_col() +
  coord_flip() +
  labs(x = "",
       y = "Total number of graduates") +
    scale_y_continuous(labels = comma_format()) + theme(legend.position = "none")
```

### What categories of majors make the most money amongst recent graduates?

```{r}
majrsProcessed %>% 
  group_by(Major_category) %>% 
  summarise(fieldMedian = median(Median)) %>% 
  mutate(Major_category = fct_reorder(Major_category,fieldMedian)) %>% 
  ggplot(aes(Major_category, fieldMedian, fill = Major_category)) +
  geom_col() +
  coord_flip() +
  labs(x = "",
       y = "median starting salary") +
    scale_y_continuous(labels = dollar_format()) + 
  theme(legend.position = "none")

```
```{r}
majrsProcessed %>% 
  mutate(Major_category = fct_reorder(Major_category,Median)) %>% 
  ggplot(aes(Major_category, Median, fill = Major_category)) + 
  geom_boxplot() + 
  expand_limits(y = 0) +
  scale_y_continuous(labels = dollar_format()) + 
  coord_flip() + 
  theme(legend.position = "none")

```
### What are the most popular majors?

```{r}
majrsProcessed %>%  
  arrange(desc(Total))%>% 
  mutate(Major = fct_reorder(Major,Total)) %>% 
  head(20) %>% 
  ggplot(aes(Major, Total, color = Major , fill = Major)) +
  geom_col() +
  scale_y_continuous(labels = comma_format()) +
  coord_flip() +
  theme(legend.position = "none")
```

### What are the least popular majors?

```{r}
majrsProcessed %>% 
   arrange(desc(Total))%>% 
  mutate(Major = fct_reorder(Major,Total)) %>% 
  tail(20) %>% 
  ggplot(aes(Major, Total, color = Major , fill = Major)) +
  geom_col() +
  scale_y_continuous(labels = comma_format()) +
  coord_flip() + 
  theme(legend.position = "none")
```

### What are the highest earning majors?

```{r}  
majrsProcessed %>% 
  mutate(Major = fct_reorder(Major, Median)) %>% 
  head(20) %>% 
ggplot(aes(Major, Median, color = Major_category)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  scale_y_continuous(labels = scales::dollar_format()) + 
  expand_limits(y=0) + 
  coord_flip()
```

### What are the lowest earning majors?

```{r}
majrsProcessed %>% 
  mutate(Major = fct_reorder(Major, Median)) %>% 
  tail(20) %>% 
  ggplot(aes(Major, Median, color = Major_category)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  scale_y_continuous(labels = dollar_format()) + 
  expand_limits(y=0) + 
  coord_flip()
```

### Correcting for sample size

```{r}
majrsProcessed %>% 
  ggplot(aes(Sample_size,Median)) + 
           geom_point() +
           geom_text(aes(label = Major), check_overlap = TRUE, vjust = 1, hjust = 1) + 
  scale_x_log10()
```

### What are the top 20 earning majors, with at least a 100 graduates responding?

```{r}  
majrsProcessed %>% 
  mutate(Major = fct_reorder(Major, Median)) %>% 
  filter(Sample_size >= 100) %>% 
  head(20) %>% 
ggplot(aes(Major, Median, color = Major_category)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  scale_y_continuous(labels = scales::dollar_format()) + 
  expand_limits(y=0) + 
  coord_flip() +
    labs( title = "The highest earning majors", subtitle = "Top 20 earning majors with at least 100 graduates responding; bars are 25th and 75th percentiles", 
    x = "",
    y = "median salary of graduates")
```

### What are the lowest 20 earning majors, with at least a 100 graduates responding?

```{r}  
majrsProcessed %>% 
  mutate(Major = fct_reorder(Major, Median)) %>% 
  filter(Sample_size >= 100) %>% 
  tail(20) %>% 
ggplot(aes(Major, Median, color = Major_category)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  scale_y_continuous(labels = scales::dollar_format()) + 
  expand_limits(y=0) + 
  coord_flip() +
    labs( title = "The lowest earning majors", subtitle = "Lowest 20 earning majors with at least 100 graduates responding; bars are 25th and 75th percentiles",
          x = "",
    y = "median salary of graduates")
```

### How does gender relate to typical earnings?

```{r}
majrsProcessed %>% 
  arrange(desc(Total)) %>% 
  mutate(Major = fct_reorder(Major, Total)) %>% 
  head(20) %>% 
  gather(Gender, Number, Men, Women) %>% 
  ggplot(aes(Major, Number, fill = Gender)) + 
  geom_col() +
  labs(title = "Women / men numbers in top 20 majors by popularity",
       x = "",
       y = "Total number of graduates") +
  scale_y_continuous(labels = comma_format()) +
  coord_flip() 
```

```{r byMajrCategory}
byMajrCategory <- majrsProcessed %>% 
  filter(!is.na(Total)) %>% 
  group_by(Major_category) %>% 
  summarise(Men = sum(Men),
            Women = sum(Women),
            Total = sum(Total),
            medianSalary = sum(Median * Sample_size) / sum(Sample_size)) %>% 
  mutate(ShareWomen = Women / Total ) %>% 
  arrange(desc(ShareWomen))
```

```{r}
library(ggrepel)

byMajrCategory %>% 
  ggplot(aes(ShareWomen,medianSalary)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  geom_text_repel(aes(label = Major_category)) +
  expand_limits(y = 0)
```

The above graph suggests a negative correlation between salary and majors women prefer. Is this correlation clearer if we plot majors (instead of major categories) against median salaries?

```{r}
majrsProcessed %>% 
  ggplot(aes(ShareWomen, Median, color = Major_category)) +
  geom_point() +
  geom_smooth(aes( group =1), method = "lm") +
  expand_limits(y = 0)

```

The above graph clearly shows a negative correlation between median salary for fresh graduates and majors attracting a larger share of women

More TBD

Most interesting question: how is this decision made? Influencing factors? : biology & behavior project
