---
title: "horror_movie_profits"
author: "ASG"
date: "6/3/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(scales)
library(lubridate)
```
### Original dataset : https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-10-23

### The original TidyTuesday github page posits the question: scary-movies-are-the-best-investment-in-hollywood---fivethirtyeight

### getting the raw data
```{r include=FALSE}
movie_profit <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/2e9bd5a67e09b14d01f616b00f7f7e0931515d24/data/2018/2018-10-23/movie_profit.csv") %>% 
select(-X1) %>% 
  mutate(release_date = mdy(release_date))
```

```{r}
movie_profit %>% 
  head()
```
As seen above, the dataset consists of ~3400 movies with some associated data like release date, genre, production budget, US / world-wide ticket sales etc.

Before we get into distinct genres, we could ask an obvious question: what have been the most profitable Hollywood movies made until 2018 (the database stops here) and what genre were they?

### creating a gross2cost column

The ratio of worldwide_gross sales divided by the production budget will give us an approximate measure of the profitability of a movie, not necessarily the final profit. We also filter out movies that made no money at all (worldwide_gross has to be > 0).

```{r include=FALSE}
movie_profit<- movie_profit %>%
  filter(worldwide_gross > 0) %>% 
  mutate(gross2cost_ratio = worldwide_gross / production_budget) %>% 
  arrange(desc(gross2cost_ratio))
```

### Top 20 all-time profitable movies
What genre were the top 20 all time top profitable movies?

```{r}
all_time_top_profitables <-movie_profit %>% 
  head(20)
```

```{r}
all_time_top_profitables %>% 
  mutate(movie = fct_reorder(movie,gross2cost_ratio)) %>% 
  ggplot(aes(movie, gross2cost_ratio, fill = genre)) + 
  geom_col() +
  coord_flip() +
  labs(title = "all-time-top 20 profitable movies",
       y = "worldwide gross sales / production budget",
       x = "")
```

Almost half (9/20) of the top 20 all-time profitable movies are horror flicks (pink bars), the remaining 11 include 6 dramas, 2 comedies, 2 adventures and 1 action movie. 

The top two movies "Paranormal Activity" and "The Blair Witch Project", both horror movies, sold tickets worth more than 400 times their production cost!

Is this statistically true? Do most of the profitable horror movies make much more money than the profitable movies from other genres? To answer that, we need to look at the statistical distribution of profitable movies by genre.

### looking at the gross sales to production cost ratio distribution

```{r}
movie_profit %>% 
  ggplot(aes(gross2cost_ratio, fill = genre)) +
  geom_histogram() +
  scale_x_log10(labels = comma_format())
```

The profits distribution above is log normally distributed and spans a range of 7 orders of magnitude. From a gross2cost ratio of 0.0001 (1e-04) to 1000 (1e+03). 

To the eye, all the distributions look superimposed and quite misleading, given the logarithmic X axis. Also, the number of horror movies that have been made is much smaller than the number of movies made in the other genres.

We could transform the distributions to cumulative and normalize the data (get rid of the numbers discrepancy), to bring out the differences in profits made by each genre.

### Normalized cumulative distribution of gross2cost ratio by genre

```{r}
movie_profit %>% 
  ggplot(aes(gross2cost_ratio, color = genre)) +
  stat_ecdf() +
  facet_wrap(facets = vars(genre)) +
  scale_x_log10(labels = comma_format()) +
  labs(title = "Cumulative density of worldwide gross receipts to production costs by genre",
       y = "cumulative density",
       x = "worldwide gross / production budget") +
  geom_hline(yintercept = c(0.5, 0.99), linetype = 3) +
  geom_vline(xintercept = c(1,10,100), linetype = 3)
```
As seen above, the horror flicks grossing in the top 25th percentile had ~10 - ~100 fold sales over costs! The closest among other genres are comedy movies with the top 25th percentile comedies grossing between ~5 to ~20 fold sales over costs. 

To examine this further, we filter out the loss making and least profitable movies and re-plot the data

```{r}
movie_profit %>% 
  arrange(desc(gross2cost_ratio)) %>% 
  filter(gross2cost_ratio > 1.001) %>% 
  ggplot(aes(gross2cost_ratio, color = genre)) +
  stat_ecdf() +
  scale_x_log10(labels = comma_format()) +
  labs(title = "Cumulative density of worldwide gross receipts to production costs by genre",
       y = "cumulative density",
       x = "worldwide gross / production budget") +
  geom_hline(yintercept = c(0.5, 0.99), linetype = 3) +
  geom_vline(xintercept = c(5,10,50,100), linetype = 3)
```

We see that the horror movie profitability distribution stands out from all other genres! The median horror flick made ~ 4 times production cost in sales (i.e ~ 400 % gross profit) compared to a median comedy movie which made ~ 2 fold sales over costs. 

The top horror flicks (99th percentile) made over 400 times production costs (an astonishing 40,000 % gross profit) compared to the top grossing comedy movies that, at best, made 30 times costs or ~ 3000 % gross profit. 

Clearly, horror movies are the best genre to invest money in!


TBD 

### Are there any distributors who have a special talent for picking profitable horror flicks? 

